### 2017-06-15 paper/paper.tex
small fix and test git on mac

\documentclass[10pt,a4paper]{article}
\RequirePackage[italian]{babel}
\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{sectsty}

\sectionfont{\fontsize{13}{15}\selectfont}
\subsectionfont{\fontsize{11}{12}\selectfont}

\author{Marco Faretra, Gabriele Marini}
\title{\textbf{Big Data Analysis on NBA \& NCAA}\\Colleges' impact on professionistic games}
\begin{document}

\maketitle
\thispagestyle{empty}
\pagestyle{empty}

\section*{RIASSUNTO}

\section{INTRODUZIONE}

\section{RISORSE}
\section{APPROCCIO}
\section{VALUTAZIONE}

\section{CONCLUSIONI}

\section{SVILUPPI FUTURI}

\end{document}